---
title: "Zhongdi's Proejct"
author: "Jiaq Yin"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  bookdown::html_document2:
    toc: no
    number_sections: yes
    fig_caption: yes
    fig_width: 7
    fig_height: 4
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
- \usepackage{amsmath}
- \usepackage{dsfont}
- \usepackage{booktabs}
- \usepackage{longtable}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage[table]{xcolor}
- \usepackage{wrapfig}
- \usepackage{colortbl}
- \usepackage{pdflscape}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=TRUE, message=FALSE, warning=FALSE, autodep=TRUE,fig.path = 'figures/', fig.pos='center')

rm(list = ls())
```

```{r library}
library(gee)
library(readr)
library(dplyr)
library(pander)
library(kableExtra)
library(data.table)
library(DT)
library(ggplot2)
```
# Data Exploration
```{r loadData}
dat = read_csv("ThresSize3x3.csv", col_names = FALSE)
names(dat) = c("measure", "age", "id", "dist")
dat = dat %>% filter(dist != 5)
dat$dist.f = as.factor(dat$dist)

dat_axis = read_csv("ThresMinorAxis3x3.csv")
names(dat_axis) = c("measure", "age", "id", "dist")
dat_axis = dat_axis %>% filter(dist != 5)
dat_axis$dist.f = as.factor(dat_axis$dist)
```

## Area
For each participants, we collected the age, and measured the size of cells multiple times with respect to different distance. In the following, we display the number of measurements, mean of size, and its standard deviation. 


```{r descriptiveEachSample}
measure_per_person = dat %>% group_by(id, dist, age) %>% 
        summarise(n = n(), mean = mean(measure), sd = sd(measure))
datatable(measure_per_person,  class="table-condensed", options = list(dom = 'tp',scrollX = TRUE))
```
In Fig.(\@ref(fig:SizeVsAge)), each point represents the mean size of one participant w.r.t different distance types. 

```{r SizeVsAge, fig.cap="Size Vs. age"}
measure_per_person$dist.f = as.factor(measure_per_person$dist)
measure_per_person %>% ggplot(aes(x = age, y = mean, color = dist.f, linetype = dist.f)) +
        geom_point() + 
        geom_smooth(aes(linetype=dist.f, color=dist.f)) + # Add a loess smoothed fit curve with confidence region 
        ylab("Mean Size") +
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        scale_color_discrete(name = "Distance") + 
        scale_linetype_discrete(name = "Distance") + 
        theme_bw()
```

### Empirical Results
Without considering the clustering issue.

```{r AreaEmp, fig.cap="Area: Empirical area vs. age"}
area_tmp = dat %>% group_by(age, dist.f) %>% summarise(n = n(), mean_area = mean(measure), sd_area = sd(measure)) %>% mutate(L = mean_area - sd_area, U = mean_area + sd_area)
area_tmp %>%
        ggplot(aes(x = age, fill=dist.f)) + 
        geom_ribbon(aes(ymin=L, ymax = U), alpha = 0.3) + 
        geom_line(aes(y = mean_area, color = dist.f)) + 
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        theme_bw()
```
From the plot, we realize there are outliers when collecting data, which are shown in the following table. The number of measurements for such errors are very small. We further delete them from our original dataset. 
```{r removeError}
# measure_per_person %>% filter(dist == 3, mean > 3000) %>%
#         kable(caption = "Errors When Collectiong Data") %>%
#         kable_styling()
# dat2: remove errors
dat2 = dat
```
Before we start regression, we check the distribution of predictors. In Fig.(\@ref(fig:countAge)), the number of measurements for distance as 3 is rather small compared with the other two types. It can causes large variation when we preform regression methods.
```{r countAge, fig.cap="The number of measurements w.r.t ages and distance"}
dat2$dist.f = as.factor(dat2$dist)
dat2 %>% group_by(age, dist.f) %>% summarise(n = n()) %>%
        ggplot(aes(x=age,y=n, fill = dist.f)) + 
        geom_bar(stat="identity",position=position_dodge()) + 
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        theme_bw()
```
In the follwowing, size varies with age w.r.t different distance.

Distance as 1
```{r Dist1MeasureVsAge, fig.cap="Distance = 1 Measure vs. Age"}
tmp1 = dat2 %>% filter(dist == 1)
tmp1 %>%
        ggplot(aes(x = age, y = measure)) + 
        geom_point() +
        theme_bw()
```
Distance as 2.
```{r Dist2MeasureVsAge, fig.cap="Distance = 2 Measure vs. Age"}
tmp2 = dat2 %>% filter(dist == 2)
tmp2 %>%
        ggplot(aes(x = age, y = measure)) + 
        geom_point() +
        theme_bw()
```
Distance as 3.

```{r Dist3MeasureVsAge, fig.cap="Distance = 3 Measure vs. Age"}
tmp3 = dat2 %>% filter(dist == 3)
tmp3 %>%
        ggplot(aes(x = age, y = measure)) + 
        geom_point() +
        theme_bw()
```
## Axis

```{r AxisdescriptiveEachSample}
measure_per_person_axis = dat_axis %>% group_by(id, dist, age) %>% 
        summarise(n = n(), mean = mean(measure), sd = sd(measure))
datatable(measure_per_person_axis,  class="table-condensed", options = list(dom = 'tp',scrollX = TRUE))
```


```{r AxisSizeVsAge, fig.cap="Axis: Size Vs. age"}
measure_per_person_axis$dist.f = as.factor(measure_per_person_axis$dist)
measure_per_person_axis %>% ggplot(aes(x = age, y = mean, color = dist.f, linetype = dist.f)) +
        geom_point() + 
        geom_smooth(aes(linetype=dist.f, color=dist.f)) + # Add a loess smoothed fit curve with confidence region 
        ylab("Mean Size") +
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        scale_color_discrete(name = "Distance") + 
        scale_linetype_discrete(name = "Distance") + 
        theme_bw()
```
### Empirical Results
Without considering the clustering issue.

```{r AxisEmp, fig.cap="Area: Empirical area vs. age"}
axis_tmp = dat_axis %>% group_by(age, dist.f) %>% summarise(n = n(), mean_axis = mean(measure), sd_area = sd(measure)) %>% mutate(L = mean_axis - sd_area, U = mean_axis + sd_area)
axis_tmp %>%
        ggplot(aes(x = age, fill=dist.f)) + 
        geom_ribbon(aes(ymin=L, ymax = U), alpha = 0.3) + 
        geom_line(aes(y = mean_axis, color = dist.f)) + 
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        scale_y_continuous(breaks = seq(0,80, 5)) + 
        theme_bw()
```

From the plot, we realize there are outliers when collecting data, which are shown in the following table. The number of measurements for such errors are very small. We further delete them from our original dataset. 
```{r AxisremoveError}
# measure_per_person_axis %>% filter(dist == 3, mean > 60) %>%
#         kable(caption = "Errors When Collectiong Data") %>%
#         kable_styling()
# dat2: remove errors
dat2_axis = dat_axis #%>% filter(id != 95 | dist != 3) %>% filter(id != 161 | dist != 3)
```

```{r AxiscountAge, fig.cap="Axis: The number of measurements w.r.t ages and distance"}
dat2_axis$dist.f = as.factor(dat2_axis$dist)
dat2_axis %>% group_by(age, dist.f) %>% summarise(n = n()) %>%
        ggplot(aes(x=age,y=n, fill = dist.f)) + 
        geom_bar(stat="identity",position=position_dodge()) + 
        scale_x_continuous(breaks = seq(15, 90, 5)) +
        theme_bw()
```

# Regression

## Area
Consider of multiple measurements for each participants, we perform General Estimated Equation (GEE). Treat each participants as one cluster. The outcome is size, and the predictors are age and distance type. Age are treated as continouse variables, and distance type as category variables.  We assign independence working matrix to each cluster. Robust standard erros are used and 95% confidnece interval is reported in the following table
```{r gee}
#geelm = gee(measure ~ age + dist.f, id = id, data = dat2, corstr = "independence", family = "gaussian")
geelm = gee(measure ~ age + dist.f, id = id, data = dat2, corstr = "independence", family = "gaussian")


se = summary(geelm)$coefficients[,c("Robust S.E.")]
gee_tab = data.frame(Coef = coef(geelm), SE = se, "L" = coef(geelm) - se * qnorm(0.975), "U" = coef(geelm) + se * qnorm(0.975), pvalue = 2 * (1 - pnorm(abs(summary(geelm)$coefficients[,5])))) 

kable(gee_tab, caption = "GEE Results with Robust S.E") %>%
        kable_styling()
```
## Axis
interval is reported in the following table
```{r gee_axis}
geelm_axis = gee(measure ~ age + dist.f, id = id, data = dat2_axis, corstr = "independence", family = "gaussian")

se = summary(geelm_axis)$coefficients[,c("Robust S.E.")]
gee_tab = data.frame(Coef = coef(geelm_axis), SE = se, "L" = coef(geelm_axis) - se * qnorm(0.975), "U" = coef(geelm_axis) + se * qnorm(0.975), pvalue = 2 * (1 - pnorm(abs(summary(geelm_axis)$coefficients[,5])))) 

kable(gee_tab, caption = "GEE Results with Robust S.E") %>%
        kable_styling()
```
# Prediction

```{r mydata}
dist_level = 4
mydata = data.frame(intercept = rep(1,dist_level*100), age = rep(1:100, dist_level))
dist.f = as.factor(rep(c(1:dist_level), each = 100))
dist_tmp = model.matrix(~ dist.f)[,-1]
colnames(dist_tmp) = paste("dist.f", c(2:dist_level), sep = "")
        
mydata = as.matrix(cbind(mydata, dist_tmp))
```
## Area
```{r AreaMydata, fig.cap="Area: prediction with 95% CI"}
mydata_y = mydata %*% as.matrix(coef(geelm))
sd_vec = summary(geelm)$coefficients[, "Robust S.E."]
var_mat = diag(sd_vec^2)
mydata_sd = sqrt(apply(mydata, 1, function(x) t(x) %*% var_mat %*% x))

L = mydata_y - 1.96 * mydata_sd
U = mydata_y + 1.96 * mydata_sd

mydata_predict = data.frame(age = rep(1:100, dist_level), dist.f = as.factor(rep(1:dist_level, each = 100)), area_size = mydata_y, L = L, U=U)

mydata_predict %>% ggplot(aes(x = age, fill=dist.f)) + 
        geom_ribbon(aes(ymin=L, ymax = U), alpha = 0.3) + 
        geom_line(aes(y = area_size, color = dist.f)) + 
        scale_x_continuous(breaks = seq(0, 100, 5)) +
        #ylim(c(0,80))+
        #scale_y_continuous(breaks = seq(0,80,5)) +
        theme_bw()
```
## Axis
```{r axisMydata, fig.cap="Axis: prediction with 95% CI"}
mydata_y = mydata %*% as.matrix(coef(geelm_axis))
sd_vec = summary(geelm_axis)$coefficients[, "Robust S.E."]
var_mat = diag(sd_vec^2)
mydata_sd = sqrt(apply(mydata, 1, function(x) t(x) %*% var_mat %*% x))

L = mydata_y - 1.96 * mydata_sd
U = mydata_y + 1.96 * mydata_sd

mydata_predict = data.frame(age = rep(1:100, dist_level), dist.f = as.factor(rep(1:dist_level, each = 100)), axis_size = mydata_y, L = L, U=U)

mydata_predict %>% ggplot(aes(x = age, fill=dist.f)) + 
        geom_ribbon(aes(ymin=L, ymax = U), alpha = 0.3) + 
        geom_line(aes(y = axis_size, color = dist.f)) + 
        scale_x_continuous(breaks = seq(0, 100, 5)) +
        #ylim(c(0,80))+
        #scale_y_continuous(breaks = seq(0,80,5)) +
        theme_bw()
```
## Diagnostics
```{r AxisfitValueVsAge, fig.cap="Fit value Vs. truth w.r.t age and distance type"}
# dat2$fittedValue = fitted(geelm)
# 
# tmp1 = dat2 %>% filter(dist == 1)
# tmp1 %>%
#         ggplot(aes(x = age, y = measure)) + 
#         geom_point() + 
#         geom_line(color = "red", data = tmp1, aes(x = age, y = fittedValue))
```
